Scheeper's Meager-NWD Game and the Menger Game

### 2014-11-27 phd/dissertation/chapter5.tex
free ultrafilters

\section{Predetermined strategies for $\gruKPGame{X}$, $\gruKLGame{X}$}

Adding knowledge of the round number to a tactic changes the characterization
from metacompact to $\sigma$-metacompact. In fact, the analogous result holds
for $0$-tactics to $0$-marks, known as predetermined strategies since they
rely only on the round number and not the moves of the opponent.

@@

\section{Compactly generated spaces}

@@

  Let $L:\omega\to\mc{K}(X)$ be a legal attack against $\sigma$, and let
  $L_{\omega\setminus n} = \bigcup_{n\leq m<\omega}L(m)$. Then as
    \[
      L_{\omega\setminus n}\cap K_n
        =
      \bigcup_{n\leq m< p}L(m) \cap \sigma(p)
    \]
  is compact for each $p<\omega$, $L_{\omega\setminus n}$ is closed.

@@

% \section{Non-locally compact spaces}

% \begin{prop}
%   If $X=\omega\cup\mathcal{F}\subset\beta\omega$ (a non-$k$ space),
%   then $K\not\prewin \gruKLGame{X}$.
% \end{prop}

>>>>

\section{Locally compact spaces and predetermined strategies}

As mentioned above, adding knowledge of the round number to a tactic changes the
characterization from metacompact to $\sigma$-metacompact. In fact, the
analogous result holds
for $0$-tactics to $0$-marks, known as predetermined strategies since they
rely only on the round number and not the moves of the opponent.

@@

\section{Compactly generated spaces and predetermined strategies}

@@

  Let $L:\omega\to\mc{K}(X)$ be a legal attack against $\sigma$, and let
  $L_{\omega\setminus n} = \bigcup_{n\leq m<\omega}L(m)$. Then as
    \[
      L_{\omega\setminus n}\cap K_p
        =
      \bigcup_{n\leq m< p}L(m) \cap \sigma(p)
    \]
  is compact for each $p<\omega$, $L_{\omega\setminus n}$ is closed.

@@

\section{Non-equivalence of $\gruKPGame{X}$, $\gruKLGame{X}$}

For $k$-spaces, it has been shown that $\gruKPGame{X}$ and $\gruKLGame{X}$
are equivalent with respect to $\pl K$'s winning predetermined strategies.
Looking at a subspace of the Stone-Cech compactification $\beta\omega$ of
$\omega$ reveals an example for which the predetermined strategies are not
equivalent.

\begin{defn}
  An \term{ultrafilter} on a cardinal $\kappa$ is a maximal filter of non-empty
  subsets of $\kappa$. For each $\alpha\in \kappa$, the ultrafilter
  $\mc F_\alpha$ containing all supersets of $\{\alpha\}$ is called a
  \term{principal ultrafilter}. All
  ultrafilters not of this form are called \term{free ultrafilters}.
\end{defn}

\begin{defn}
  The \term{Stone-Cech compactification} of a cardinal
  $\kappa$ is the space $\beta\kappa$ consisting
  of all ultrafilters on $\kappa$, with open sets of the form
  $U_S=\{\mc F\in\beta\kappa : S\in\mc F\}$ for $S\subseteq \kappa$.
\end{defn}

From these definitions it is easily verified that principal ultrafilters
are isolated, so $\kappa$ with the discrete topology may be viewed as
a dense open subspace of $\beta\kappa$. $\beta\kappa$ is also compact, so
$\pl K\ktactwin{0}\gruKLGame{\beta\kappa}$; of greater interest is the subspace
of $\beta\omega$ consisting of all principal ultrafilters and a single free
ultrafilter $\mc F$, denoted $\omega\cup\{\mc F\}$.

\begin{lem}
  All compact subsets of $\omega\cup\{\mc F\}\subset\beta\omega$ are finite.
  In particular, the difference of compact sets in $\omega\cup\{\mc F\}$
  is compact.
\end{lem}

\begin{proof}
  Let $I=\{n_i:i<\omega\}\cup\{\mc F\}$ be infinite.
  Then $\{U_{\omega\setminus\{n_i:i\geq j\}}:j<\omega\}$ is an open cover of
  $I\cup\{\mc F\}$ with no finite subcover.
\end{proof}

\begin{thm}
  $\pl K\not\prewin \gruKLGame{\omega\cup\{\mc F\}}$ for any free
  ultrafilter $\mc F$.
\end{thm}

\begin{proof}
  Let $\sigma$ be a predetermined strategy for $\pl K$, and define the
  counterattack $H:\omega\to\mc K(X)$ by
  $H(n)=(n\cup\sigma(n+1))\setminus\sigma(n)$. Then for any neighborhood
  $U_S$ of $\mc F$, $S$ is infinite, and since
  $\bigcup_{n<\omega} H(n)\supseteq\omega\setminus\sigma(0)$, $U_S$ meets
  infinitely many of the finite $H(n)$. Thus $\sigma$ is not a winning
  predetermined strategy.
\end{proof}

\begin{thm}
  There exists a free ultrafilter $\mc F$ such that
  $\pl K\prewin \gruKPGame{\omega\cup\{\mc F\}}$.
\end{thm}

\begin{proof}
  Let $\mc F$ be any free ultrafilter, and
  define the predetermined strategy $\sigma$ by
  $\sigma(n)=n^2\cup\{\mc F\}$.

  Consider the set of all legal attacks $A\subseteq\omega^\omega$ by
  $\pl P$ against $\sigma$. For $n<\omega$, note that for any $f\in A$,
  $f$ maps only $n$ points into $n^2$, so the range of $f$ is coinfinite.
  Then $\mc I=\{\ran f : f\in A\}$ is an ideal, and if
  $\mc F\supseteq\{\omega\setminus I: I\in\mc I\}$, then $\sigma$ is a
  winning predetermined strategy.
\end{proof}
